---
title: "Hyperspectral imaging"
author: "Edmund Hart"
date: "August 6, 2014"
output: html_document
---

Reading data

```{r Reading in data}
library(raster)
library(rhdf5)
f <- '/Users/thart/ESAData_2014/Part2_Spectrometer/HDF5File/SJER_140123_chip.h5'
h5ls(f,all=T)
```

Grab metadata
```{r get spatial info and map info}
spinfo <- h5metadata(f,"spatialInfo",11)
```


Slice up bands make a quick plot
```{r slicing}
b425<- h5read(f,"Reflectance",index=list(1:477,1:502,425))
## Convert from array to matrix
b425 <- t(b425[,,1])
b425[b425 > 14999] <- NA
image(log(t(b425)))
```

Ok, now let's create a meaninful raster from our data
```{r create raster}
b425r <- raster(t(b425))

## Now we need to grab the extent in a bounding box.  Luckily this is in the metadata,  We'll need it in the form of LL lon, UR lon, LL lat, UR lat

ex <- sort(unlist(spinfo[2:5]))
e <- extent(ex)
extent(b425r) <- e
plot(b425r)
```

Awesome, now that we've plotted one band, try plotting your own maps with some different bands.


Now, let's try and generate a full color image with the RGB bands.  But before we start, let's write a function to handle some of the basic cleaning we did earlier, that way we can bulk process bands.


```{r RGB}
# f: the hdf file
# band: the band you want to grab
# returns: a cleaned up HDF5 reflectance file
getBandMat <- function(f, band){
  out<- h5read(f,"Reflectance",index=list(1:477,1:502,band))
  ## Convert from array to matrix
  out <- t(out[,,1])
  out[out > 14000] <- NA
  return(out)
}
band2rast <- function(f,band){
  out <-  raster(getBandMat(f,band),crs="+zone=11N +ellps=WGS84 +datum=WGS84 +proj=longlat")
  ex <- sort(unlist(spinfo[2:5]))
  e <- extent(ex)
  extent(out) <- e
  return(out)
}

stackList <- function(rastList){
  ## Creates a stack of rasters from a list of rasters
  masterRaster <- stack(rastList[[1]])
  for(i in 2:length(rastList)){
    masterRaster<-  addLayer(masterRaster,rastList[[i]])
  }
  return(masterRaster)
}

rgb <- list(58,34,19)
rgb_rast <- lapply(rgb,band2rast, f = f)
## Add the names of the bands so we can easily keep track of the bands in the list
names(rgb_rast) <- as.character(rgb)
### Check with a plot
plot(rgb_rast[[1]])
rgb_stack <- stackList(rgb_rast)
plot(rgb_stack)
plotRGB(rgb_stack,r=1,g=2,b=3, scale=255, stretch = "Lin")
```

We can also plot our image on a map of the US

```{r mapping}
library(maps)
map(database="state",region="california")
points(spinfo$LL_lat~spinfo$LL_lon,pch = 15)
### Add raster
```

Now the fun stuff!  Let's create NDVI
```{r ndvi}
ndvi_bands <- c(3,4)
ndvi_rast <- lapply(ndvi_bands,band2rast, f = f)
ndvi_stack <- stackList(ndvi_rast)
NDVI <- function(x) {
  (x[,2]-x[,1])/(x[,2]+x[,1])
}
ndvi_calc <- calc(ndvi_stack,NDVI)
plot(ndvi_calc)
```



  

